Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-10-11 ms.tex
reformated for nature communications submission

% Define document class
\documentclass[manuscript]{aastex63}
\usepackage{tabularx}
\usepackage{amsmath}

@@

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were still sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. We assume that prebiotic chemistry required a post-impact reducing atmosphere. Then, using statistics for the impact history on early Earth and the minimum impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis (i.e., the origin of life) is favored early in the Hadean eon. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred in a $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainty. Previous claims of a far narrower window for the origin of life are unsupported. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is favorable in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for biopoiesis on rocky planets elsewhere which will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

We assume that the the majority of the size-frequency distribution of impactors is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. Data for the frequency of main-belt asteroids only extends down to about 1 km diameter ($1.3 \times 10^{12} $ kg). To extend the distribution to smaller objects, we use the observed 1400 ratio between the frequency of $> 1$ km and $> 20$ km craters on the Moon following \citet{Morbidelli_2018}. Crater scaling relations suggest that 1 km and 20 km craters on the Moon corresponds to 50 m and 1 km asteroids, respectively \citep{Morbidelli_2018}. The largest main belt asteroid is $\sim 1000$ km, so we must extrapolate to larger impactors. Through the main text, we use the same extrapolation as \citet{Marchi_2014}, which extends the distribution to big impacts with a slope $d (\ln S_0)/d (\ln m) = - 0.415$ (Figure \ref{fig:sfd_and_flux}). In Appendix \ref{sec:append_sfd}, we show that alternatively extrapolating with the red dashed line in Figure \ref{fig:sfd_and_flux}a (with slope $d (\ln S_0)/d (\ln m) = - 1.0$) does not significantly change our overall conclusions. Finally, we normalize the size-frequency distribution to the impact mass required to make a 1 km crater on the Moon (50 m object, $m_0 = 1.64 \times 10^{8}$ kg). Figure \ref{fig:sfd_and_flux}a shows the resulting size-frequency distribution.

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution (SFD) of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and, in the main text, log-log extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. In Appendix \ref{sec:append_sfd} we also consider the red dashed extrapolation to $\gtrsim 10^{21}$ kg impacts. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Appendix Figure \ref{fig:velocity_distribution} shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close-approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10$ km). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@

Figure \ref{fig:example_impact_histories} illustrates three of 5000 impact histories computed with our Monte-Carlo approach (Section \ref{sec:methods}). In Figure \ref{fig:example_impact_histories}a, a Moneta-sized impact occurs at $\sim 4.43$ Ga delivering nearly all of Earth's mantle HSEs. The impact should also chemically reduce the Hadean atmosphere creating conditions favorable for prebiotic chemistry. However, in this case, Moneta cannot have given rise to extant life on Earth because Moneta is followed by an ocean-vaporizing collision at $\sim 4.37$ Ga which should destroy any existing biosphere. Based on smoothed-particle hydrodynamics impact simulations, we nominally assume that 10\% of an impact's kinetic energy goes into vaporizing the ocean \citep[Appendix \ref{sec:append_vap},][]{Citron_2022}. Vaporizing the ocean requires $5 \times 10^{27}$ J, so a $> 5 \times 10^{28}$ J impact should deliver sufficient energy under our assumptions. Appendix \ref{sec:append_vap} considers alternative criteria for ocean-vaporization. In Figure \ref{fig:example_impact_histories}a, the last ocean-vaporizing impact is only $\sim 10^{20}$ kg (360 km) but has a large impact velocity of $32$ km/s giving it $5.12 \times 10^{28}$ J of kinetic energy.

@@

A caveat to our results is that ocean-vaporization by an impact is likely a complicated function of impact mass, velocity, and incident angle, which we do not fully account for (Appendix \ref{sec:append_vap}). Throughout Section \ref{sec:results}, we have assumed that 10\% of an impactor's kinetic energy heats the planetary surface ($f_\mathrm{E,vap} = 0.1$), leading to a required $> 5 \times 10^{28}$ J collision to vaporize an ocean. Appendix Figure \ref{fig:energy_for_ocean_vap} uses simulations of impacts from \citet{Citron_2022} to show that $f_\mathrm{E,vap} = 0.1$ is a reasonable assumption, but values between 2.5\% and 25\% are also possible and $f_\mathrm{E,vap}$ likely also depends on impact angle.

To determine the sensitivity of our results to $f_\mathrm{E,vap}$, we recomputed the likelihood and timing of an impact that starts life in Appendix \ref{sec:append_vap} using $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$. We find that the probability of a life-starting impact is somewhat sensitive to $f_\mathrm{E,vap}$, while the timing does not depend strongly on $f_\mathrm{E,vap}$. Overall, this uncertainty does not change our qualitative conclusion that an origin of life impact on the early Earth was not a fluke. Even in our worst-case scenario (i.e., \citet{Wogan_2023} pessimistic case with $f_\mathrm{E,vap} = 0.25$) a life-starting impact occurs in 20\% of impact realizations (Appendix Figure \ref{fig:probabilities_of_impacts_sens}). Additionally, uncertainty in $f_\mathrm{E,vap}$ does not change our general conclusion that the most likely timing for the origin of life was $\sim 4.35$ Ga, with 95\% uncertainty spanning the Hadean eon (Appendix Figure \ref{fig:timing_sensitivity}). However, a more complete understanding of $f_\mathrm{E,vap}$ as a function of impact angle and other similar parameters may change this result.

@@

\section{Conclusions}

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life that requires early ribonucleobase synthesis. We use the results of \citet{Wogan_2023}, who find that significant origin of life precursor molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by an ocean-vaporizing collision that sterilizes the planet.

@@

Our simulations of Earth's impact history do not always result in a bombardment leading to an origin of life. There are some impact histories where a collision of sufficient mass to produce prebiotic molecules does not occur, or alternatively, does occur but primitive life is subsequently destroyed by an ocean-vaporizing impact that does not rekindle the biosphere. With our nominal assumptions, a life-starting impact occurs 92\% or 41\% of the time when we assume a $> 4 \times 10^{20}$ kg or $> 5 \times 10^{21}$ kg impact without later ocean-vaporization is required to start life, respectively. In our worst-case scenario, which assumes that impacts can more easily vaporize the ocean, a $> 5 \times 10^{21}$ kg impact without subsequent planet sterilization occurs in 20\% of simulated impact histories (Appendix Figure \ref{fig:probabilities_of_impacts_sens}).

@@

@@

\appendix

\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figures/velocity_distribution.pdf}
  \caption{Our assumed probability distribution for the velocity of impacts derived from modern observations of asteroid close approaches \citep{Park_2023}.}
  \label{fig:velocity_distribution}
\end{figure}

\section{Size-frequency distribution sensitivity test} \label{sec:append_sfd}

We assume that the size-frequency distribution of objects that struck the early Earth is similar to the main-belt asteroids. A problem with this approach is that the main-belt asteroids only contain objects up to $\sim$ 1000 km diameter, yet the early Earth is expected to have experienced impacts larger than this \citep{Marchi_2014}. Therefore, we must extrapolate the size-frequency distribution above $\sim$ 1000 km to larger objects. Throughout the main text we use the same extrapolation as \citet{Marchi_2014} (the black dashed line in Figure \ref{fig:sfd_and_flux}), who extends the size-frequency distribution with a log-log slope of $d (\ln S_0)/d (\ln m) = - 0.415$.

To test the sensitivity of our results to this chosen extrapolation, we re-did our Monte-Carlo analysis using a log-log slope of $d (\ln S_0)/d (\ln m) = - 1$ (the red dashed line in Figure \ref{fig:sfd_and_flux}) for impactors bigger than $\sim$ 1000 km diameter. Figures \ref{fig:probabilities_of_impacts_extrapolation_sens} and \ref{fig:timing_extrapolation_sensitivity} show how our results change when adopting this alternative extrapolation. Overall, our results are qualitatively unchanged for the different extrapolation slope ($d (\ln S_0)/d (\ln m)$).

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts_extrapolation_sens.pdf}
  \caption{Similar to Figure \ref{fig:probabilities_of_impacts}, except we consider different extrapolations of the size-frequency distribution (SFD) for impactors larger than 1000 km diameter. The line labeled $d (\ln S_0)/d (\ln m) = - 0.415$ is the black dashed extrapolation in Figure \ref{fig:sfd_and_flux}a. The $d (\ln S_0)/d (\ln m) = - 1.0$ case is shown by the red dashed extrapolation in Figure \ref{fig:sfd_and_flux}a.}
  \label{fig:probabilities_of_impacts_extrapolation_sens}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{figures/timing_extrapolation_sensitivity.pdf}
  \caption{Similar to Figure \ref{fig:time_and_mass}, except we consider different extrapolations of the size-frequency distribution (SFD) for impactors larger than 1000 km diameter. The line labeled $d (\ln S_0)/d (\ln m) = - 0.415$ is the black dashed extrapolation in Figure \ref{fig:sfd_and_flux}a. The $d (\ln S_0)/d (\ln m) = - 1.0$ case is shown by the red dashed extrapolation in Figure \ref{fig:sfd_and_flux}a.}
  \label{fig:timing_extrapolation_sensitivity}
\end{figure}

\section{Impact energy required for ocean vaporization} \label{sec:append_vap}

Recently, \citet{Citron_2022} performed smoothed-particle hydrodynamics impact simulations over a range of impact angles, masses and velocities to estimate the impact properties that can vaporize an ocean. Figure \ref{fig:energy_for_ocean_vap}a shows their simulation results for the change in the atmosphere's and ocean's internal energy ($\Delta \mathrm{IE}_\mathrm{atmo}$) as a function of impact kinetic energy ($\mathrm{E}_\mathrm{imp}$), along with log-log extrapolations for each impact angle. Figure \ref{fig:energy_for_ocean_vap} gives the same information, but the y-axis is instead the fraction of the impactor's energy that heats the atmosphere and ocean over the course of their simulations. For the most probable incident impact angle of $45^\circ$ about $\sim 10\%$ of the impactors kinetic energy heats the atmosphere ($f_\mathrm{E,vap} = 0.1$), so we nominally adopt this value in the main text to determine which collisions vaporize the ocean.

Energy delivery to the atmosphere/ocean appears to depend on impact angle, mass and velocity (Figure \ref{fig:energy_for_ocean_vap}). Additionally, all of the \citet{Citron_2022} simulations are far more massive than the minimum threshold for ocean-vaporization, so we must rely on extrapolations. Therefore, our assumption of a constant $f_\mathrm{E,vap} = 0.1$ is an over-simplification. To remedy this shortcoming, we considered building a parameterization for $f_\mathrm{E,vap}$ that depends on multiple impact properties (e.g., impact angle) using the \citet{Citron_2022} simulation results, but this was unsuccessful because their calculations do not consider a wide enough parameter space.

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/energy_for_ocean_vap.pdf}
  \caption{\citet{Citron_2022} smoothed-particle hydrodynamics simulations that show how much impactor kinetic energy heats the atmosphere and ocean. Different colors indicated various incident impact angles. (a) shows the simulated change in internal energy of the atmosphere and ocean as a function of the impactor energy, with linear log-log extrapolations for each impact angle. The dotted horizontal line at $5 \times 10^{27}$ J indicates the energy needed to vaporize an ocean \citep{Sleep_1989}. (b) contains the same information as (a), except the y-axis is the fraction of impactor energy that heats the atmosphere and ocean.}
  \label{fig:energy_for_ocean_vap}
\end{figure}

The evaluate the sensitivity of our results to an assumed constant $f_\mathrm{E,vap} = 0.1$, we recomputed the likelihood and timing of a life-starting impact using $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ (Figure \ref{fig:probabilities_of_impacts_sens} and \ref{fig:timing_sensitivity}). We chose these values because they are reasonable lower and upper bounds based on the \citet{Citron_2022} simulations (Figure \ref{fig:energy_for_ocean_vap}b). The probability of an impact that produces substantial prebiotic molecules without later ocean vaporization decreases with increasing $f_\mathrm{E,vap}$ (Figure \ref{fig:probabilities_of_impacts_sens}). For example, for $f_\mathrm{E,vap} = 0.1$, the probability of a $>4 \times 10^{20}$ kg impact (\citet{Wogan_2023} optimistic case) without subsequent ocean vaporization is 92\%. For $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$ the probabilities are instead 99.9\% and 60\%, respectively.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/probabilities_of_impacts_sens.pdf}
  \caption{Similar to Figure \ref{fig:probabilities_of_impacts}, except we consider different values for the fraction of impactor energy that heats the ocean.}
  \label{fig:probabilities_of_impacts_sens}
\end{figure}

Figure \ref{fig:timing_sensitivity} shows the timing of the last impact to make conditions favorable for biopoiesis that does not experience subsequent ocean vaporization for $f_\mathrm{E,vap}$ values of 0.025, 0.1 (nominal value), and 0.25. As in the main text, we consider the \citet{Wogan_2023} optimistic (Figure \ref{fig:timing_sensitivity}a) and pessimistic (Figure \ref{fig:timing_sensitivity}b) minimum impactor masses to produce significant prebiotic molecules. In Figure \ref{fig:timing_sensitivity}a, the timing of the last life-starting impact is relatively insensitive to $f_\mathrm{E,vap}$. In the \citet{Wogan_2023} pessimistic case (Figure \ref{fig:timing_sensitivity}b), an origin of life is preferred later in the Hadean for larger values of $f_\mathrm{E,vap}$.

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{figures/timing_sensitivity.pdf}
  \caption{Similar to Figure \ref{fig:time_and_mass}, except we consider different values for the fraction of impactor energy that heats the ocean.}
  \label{fig:timing_sensitivity}
\end{figure}

Overall, uncertainty in the impact properties required to vaporize the ocean has a small effect on our qualitative conclusions. Regardless of $f_\mathrm{E,vap}$, the origin of life on Earth from an impact does not appear to be a fluke (Figure \ref{fig:probabilities_of_impacts_sens}) and biopoiesis is preferred early in the Hadean with uncertainty spanning the entire eon (Figure \ref{fig:timing_sensitivity}). However, it is conceivable that a more complete understanding $f_\mathrm{E,vap}$ as a function of impact angle and other impact parameters could change these results.

>>>>

% Define document class
\documentclass[linenumbers]{aastex63}
\usepackage{tabularx}
\usepackage{amsmath}

@@

\begin{abstract}
  Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. We assume that prebiotic chemistry required a post-impact reducing atmosphere. Then, using statistics for the impact history on Earth and the minimum impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis is favored early in the Hadean. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred between 4.45 and 3.9 Ga within 95\% uncertainty. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is favorable in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for biopoiesis on rocky exoplanets which will have experienced an early episode of impact bombardment given how planets form.
\end{abstract}

@@

@@

Figure \ref{fig:example_impact_histories} illustrates three of 5000 impact histories computed with our Monte-Carlo approach (Section \ref{sec:methods}). In Figure \ref{fig:example_impact_histories}a, a Moneta-sized impact occurs at $\sim 4.43$ Ga delivering nearly all of Earth's mantle HSEs. The impact should also chemically reduce the Hadean atmosphere creating conditions favorable for prebiotic chemistry. However, in this case, Moneta cannot have given rise to extant life on Earth because Moneta is followed by an ocean-vaporizing collision at $\sim 4.37$ Ga which should destroy any existing biosphere. Based on smoothed-particle hydrodynamics impact simulations, we nominally assume that 10\% of an impact's kinetic energy goes into vaporizing the ocean \citep{Citron_2022}. Vaporizing the ocean requires $5 \times 10^{27}$ J, so a $> 5 \times 10^{28}$ J impact should deliver sufficient energy under our assumptions. The Supplementary Materials considers alternative criteria for ocean-vaporization. In Figure \ref{fig:example_impact_histories}a, the last ocean-vaporizing impact is only $\sim 10^{20}$ kg (360 km) but has a large impact velocity of $32$ km/s giving it $5.12 \times 10^{28}$ J of kinetic energy.

@@

A caveat to our results is that ocean-vaporization by an impact is likely a complicated function of impact mass, velocity, and incident angle, which we do not fully account for (Supplementary Materials). Throughout Section \ref{sec:results}, we have assumed that 10\% of an impactor's kinetic energy heats the planetary surface ($f_\mathrm{E,vap} = 0.1$), leading to a required $> 5 \times 10^{28}$ J collision to vaporize an ocean. Supplementary Materials Figure S4 uses simulations of impacts from \citet{Citron_2022} to show that $f_\mathrm{E,vap} = 0.1$ is a reasonable assumption, but values between 2.5\% and 25\% are also possible and $f_\mathrm{E,vap}$ likely also depends on impact angle.

To determine the sensitivity of our results to $f_\mathrm{E,vap}$, we recomputed the likelihood and timing of an impact that starts life in Supplementary Materials using $f_\mathrm{E,vap} = 0.025$ and $f_\mathrm{E,vap} = 0.25$. We find that the probability of a life-starting impact is somewhat sensitive to $f_\mathrm{E,vap}$, while the timing does not depend strongly on $f_\mathrm{E,vap}$. Overall, this uncertainty does not change our qualitative conclusion that an origin of life impact on the early Earth was not a fluke. Even in our worst-case scenario (i.e., \citet{Wogan_2023} pessimistic case with $f_\mathrm{E,vap} = 0.25$) a life-starting impact occurs in 20\% of impact realizations (Supplementary Materials Figure S5). Additionally, uncertainty in $f_\mathrm{E,vap}$ does not change our general conclusion that the most likely timing for the origin of life was $\sim 4.35$ Ga, with 95\% uncertainty spanning the Hadean eon (Supplementary Materials Figure S6). However, a more complete understanding of $f_\mathrm{E,vap}$ as a function of impact angle and other similar parameters may change this result.

@@

% \section{Conclusions}

In summary, we use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life that requires early ribonucleobase synthesis. We use the results of \citet{Wogan_2023}, who find that significant origin of life precursor molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by an ocean-vaporizing collision that sterilizes the planet.

@@

Our simulations of Earth's impact history do not always result in a bombardment leading to an origin of life. There are some impact histories where a collision of sufficient mass to produce prebiotic molecules does not occur, or alternatively, does occur but primitive life is subsequently destroyed by an ocean-vaporizing impact that does not rekindle the biosphere. With our nominal assumptions, a life-starting impact occurs 92\% or 41\% of the time when we assume a $> 4 \times 10^{20}$ kg or $> 5 \times 10^{21}$ kg impact without later ocean-vaporization is required to start life, respectively. In our worst-case scenario, which assumes that impacts can more easily vaporize the ocean, a $> 5 \times 10^{21}$ kg impact without subsequent planet sterilization occurs in 20\% of simulated impact histories (Supplementary Materials Figure S5).

@@

We assume that the the majority of the size-frequency distribution of impactors is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. Data for the frequency of main-belt asteroids only extends down to about 1 km diameter ($1.3 \times 10^{12} $ kg). To extend the distribution to smaller objects, we use the observed 1400 ratio between the frequency of $> 1$ km and $> 20$ km craters on the Moon following \citet{Morbidelli_2018}. Crater scaling relations suggest that 1 km and 20 km craters on the Moon corresponds to 50 m and 1 km asteroids, respectively \citep{Morbidelli_2018}. The largest main belt asteroid is $\sim 1000$ km, so we must extrapolate to larger impactors. Through the main text, we use the same extrapolation as \citet{Marchi_2014}, which extends the distribution to big impacts with a slope $d (\ln S_0)/d (\ln m) = - 0.415$ (Figure \ref{fig:sfd_and_flux}). In the Supplementary Materials, we show that alternatively extrapolating with the red dashed line in Figure \ref{fig:sfd_and_flux}a (with slope $d (\ln S_0)/d (\ln m) = - 1.0$) does not significantly change our overall conclusions. Finally, we normalize the size-frequency distribution to the impact mass required to make a 1 km crater on the Moon (50 m object, $m_0 = 1.64 \times 10^{8}$ kg). Figure \ref{fig:sfd_and_flux}a shows the resulting size-frequency distribution.

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/SFD_and_flux.pdf}
  \caption{Our assumed (a) size-frequency distribution (SFD) of impactors and (b) lunar cratering record. Most of the size-frequency distribution is identical to the main-belt asteroids \citep[Extended Data Figure 1 in][]{Marchi_2014}. We extrapolate to objects $\lesssim 10^{11}$ kg using the observed frequency ratio of $> 1$ km and $> 20$ km lunar craters following \citet{Morbidelli_2018}, and, in the main text, log-log extrapolate to asteroids $\gtrsim 10^{21}$ kg following \citet{Marchi_2014}. In the Supplementary Materials we also consider the red dashed extrapolation to $\gtrsim 10^{21}$ kg impacts. The lunar cratering record is the red line in Figure 5 of \citet{Morbidelli_2018}.}
  \label{fig:sfd_and_flux}
\end{figure}

Our final step is to assign impact velocities to each collision in the many sampled impact histories. Supplementary Materials Figure S1 shows Earth's impact velocity distribution. We created this distribution using the JPL database of close approaches to Earth by small bodies \citep{Park_2023}, considering all close-approach asteroids within 0.05 AU to Earth. The database gives the approach velocity of each asteroid far from Earth, so we compute the impact velocity by accounting for Earth's gravitational potential energy. Here, we are assuming that the impact velocity distribution for small bodies (e.g. $< 10$ km) is identical to the velocity distribution for bigger asteroids (e.g. $> 10$ km). This may not be the case because small bodies are more strongly influenced by processes like the Yarkovsky effect \citep{Bottke_2006}, where an asteroid's orbit is altered by anisotropic thermal emission.

@@
